---
title: "Assignment 6"
subtitle: "STAT 201"
author: "Matthew Oh"

format:
  html:
    toc: true
    embed-resources: true

execute:
  message: false
  warning: false
---

## Instructions {-}

1. **Write your name** on the assignment

2. Write your code in the *Code* cells of the **template provided** to write solutions for the assignment. **Do not open a new qmd**, and work from scratch. Ensure that the solution is written neatly enough to understand and grade.

3. You may talk to a friend, discuss the questions and potential directions for solving them. However, you need to write your own solutions and code separately, and not as a group activity. Do not use AI to solve the problems.

4. There are point deductions if your code is not clean and organized.

5. Only print what is asked for. Highly inefficient code will be deducted for because it demonstrates lack of understanding.

6. Render as an HTML and **submit BOTH the .qmd and HTML file**.

The grader cannot grade what they cannot see! Failure to render your document after completing all questions will lead to deductions.


## Question 1 (3 points)

### Part a

Create a vector called `bears` which contains the values `20, 17, 10, 28, 13, 30, 13, 17, 16, 26, 12, 28, 17, 27`.

Create a vector called `opponent` which contains the values `27, 41, 31, 20, 19, 12, 30, 24, 13, 31, 10, 13, 20, 16`.

Note: each element represents the amount of points scored in a particular football game.

No need to print any output here.

```{r}

bears <- c(20, 17, 10, 28, 13, 30, 13, 17, 16, 26, 12, 28, 17, 27)
opponent <- c(27, 41, 31, 20, 19, 12, 30, 24, 13, 31, 10, 13, 20, 16)

```


### Part b

Insert the value `40` as the 5th element in `bears`.

Insert the value `38` as the 1st element in `opponent`.

Print both vectors to prove you were successful.

```{r}

bears <- append(bears, 40, after=4)
opponent <- c(38,opponent)
print(bears)
print(opponent)


```


### Part c

Determine the largest win spread for the Bears (ie: spread is the difference between bear's score and opponent's score) and print the following statement:

"The largest win spread is {}, when the Bears scored {} points and the Opponent scored {} points."

Note: this needs to still run correctly even if the elements changed or were re-ordered.

```{r}

largest_spread <- (bears-opponent) #Finds the difference in scores
max_spread = max(largest_spread) #Returns the largest difference in scores where the bears win

max_index <- which.max(largest_spread) #Finds the index where this max score is achieved
bears_max <- bears[max_index] #Returns the score for the bears
opponent_max <- opponent[max_index] #Returns the score for the opponents

sprintf("The largest win spread is %s, when the Bears scored %s points and the Opponent scored %s points.", max_spread, bears_max, opponent_max)
```


## Question 2 (7 points)

### Part a

Create a matrix `game_score` by utilizing the vectors `bears` and `opponent` from Question 1 where the first column represents `bears` and the second column represents `opponent`. 

Name the columns `bears` and `opponent` respectively.

Print `game_score` to prove this was successful.

```{r}

game_score <- cbind(bears, opponent)

print(game_score)


```


### Part b

The `game_score` is missing some games. Add on a matrix that contains bears scores of `37, 9` and opponent scores of `17, 17`, respectively. Print the new `game_score` matrix.

```{r}

small_bear <- c(37,9) #create a vector that contains the new scores that we want added

small_opponent <- c(17,17)

small_score <- cbind(small_bear,small_opponent) #Create a smaller matrix that contains the bears scores and the opponents scores

game_score <- rbind(game_score, small_score) #Combine them

print(game_score) #Print


```


### Part c

Iterate through `game_score` and print "Win" if the bears won and "Loss" if the bears lost.

```{r}
for (i in 1:nrow(game_score)){  
  if (game_score[i, "bears"] > game_score[i,"opponent"]){
    print("Win")
  }else{
    print("Loss")
  }
}



```


### Part d

Use `game_score` and an **inequality** to check if the bears won each game. Store this as a vector called `win` and print the vector `win`.

ie: Since 20 is less than 38 the first element of `win` should be FALSE.

```{r}

win <- game_score[,1] > game_score[,2]
print(win)


```


### Part e

Add the vector `win` as a third column to `game_score`. Print `game_score`.

```{r}

game_score <- cbind(game_score,win)
print(game_score)

```


### Part f

How many games did the Bears lose before their first win (games are in order)? Many ways to accomplish this BUT you must utilize `game_score` to do so. 

Clarification: do not directly reference any of the individual vectors from previous question parts, this must use the matrix `game_score`. 

The code must be capable of running correctly even if the values changed.

```{r}

count <- 0 

for (i in 1:nrow(game_score)){   #Count will only go up if the win value is equal to 0 
  if (game_score[i, "win"] == 0){
    count <- count + 1
  }else{ #Once the bears reach their first win, breaks out of for loop and stops counting
    break
  }
}
print(count)

```


## Question 3 (2 points)

Write a function called `number_of_factors` that takes an integer and returns how many factors the number has. Run your function with `8` to check if it works.

```{r}
number_of_factors <- function(integer) {
  count <- 0
  for (i in 1:integer) {
    if (integer %% i == 0) {
      count <- count + 1
    }
  }
  return(count)  # This should be outside the loop
}

print(number_of_factors(8))
  
  
```


## Question 4 (3 points)

Write a function that calculates the area of a rectangle. The function should have an input for length (in inches) and width (in inches). Set a default value for both length and width to be 48.

The function should have 2 outputs: the area in terms of square inches and the area in terms of square feet. There are 12 inches in 1 foot.

Run your function with a length of 102 and width of 60 and store the output as a variable. Then print the following:

"The area in square inches is {}."

"The area in square feet is {}."

```{r}

rectangle_area <- function(length=48, width=48){
  inch_area <- length*width
  feet_area <- inch_area/144
print(paste("The area in square inches is ", inch_area, ".", sep = ""))
print(paste("The area in square feet is ", feet_area, ".", sep = ""))
}

find_area <- rectangle_area(102,60)


```


## Question 5 (5 points)

Write a function that takes a single string vector as an input. Have the function return only the word (or words) with the most letters. 

If there are multiple words with the most letters, then return all words with the most letters **as a vector**.

Clarification: This is NOT returning a list that contains each word individually. It is returning a vector containing all words with the most letters.

Test the function using the vector `fruits` created below.

```{r}
fruits <- c("apple", "orange", "banana", 
            "pear", "grapes", "kiwi")
```


```{r}
most_letters <- function(string_vector){
  character_count <- c()
  for (i in string_vector){
    character_count <- c(character_count,nchar(i)) #Adds the length of each word's length to a vector
  } 
  max_character <- max(character_count) #Gets the maximum value in the vector
  return(string_vector[character_count == max_character]) 
  
}

print(most_letters(fruits))

```


## Question 6 (4 points)

Write a function that checks if a value is in a vector.
The function should take both a value and vector as inputs and return either TRUE or FALSE.
  
Run your function using the variables `value` and `check` created below.
  
```{r}
value <- 2
check <- seq(1, 15, 3)

check_value <- function(vector,value){
 equivalent <- vector[vector == value]
  return (length(equivalent) != 0) #If the returned val
}
check_value(check,value)

```

After running the function, depending on what the function returns, print either:

  - "{} is in the vector."
  - "{} is not in the vector.
  
```{r}

in_vector <- check_value(check,value)



if (in_vector == TRUE){
  cat(value,"is in the vector.")
}else{
  cat(value,"is not in the vector.")
}
```


## Question 7 (6 points)

Below is a matrix of random numbers.

```{r}
# Don't change this! A seed keeps results consistent.
set.seed(123)
x <- matrix(sample(-50:50, 20), nrow = 5)
```

### Part a

Write a function that takes a matrix as an input, converts all negative numbers to NA, and returns the new matrix. Test your function using the matrix `x` that was created for you.

```{r}
replace_negative <- function(matrix){
  matrix[matrix<0] <- NA
  return(matrix)
}
replace_negative(x)

```

### Part b

Write a function that takes a matrix as an input and converts all even numbers to 1 and odd numbers to 0. The function should return the new matrix. Test your function using the matrix `x` that was created for you.

```{r}
convert_even_odd <-function(matrix){
  for (i in 1: nrow(matrix)){
    for (j in 1:ncol(matrix)){
      if (matrix[i,j]%%2 == 0){
        matrix[i,j] <- 1
      }else{
        matrix[i,j] <- 0 
      }
    }
  }
  return(matrix)
}

convert_even_odd(x)

```

